---
title: "Data in the Wild--Analyses"
format: gfm
execute:
  message: false
editor: visual
---

This document contains the analyses for the Data in the Wild course.

## Packages and Data

Load the packages used in analyses.

```{r}
#| message: false
library(tidyverse)
library(lme4)
library(lmerTest)
library(broom.mixed)
library(effectsize)
```

Read in the cleaned dataset.

```{r}
#| message: false
data <- read_csv("../data_clean/data_cleaned.csv")

head(data)
```

## Check Distributions

Look at the distribution of responses by question.

```{r}
data %>% 
  # reshape the data for plotting
  pivot_longer(starts_with("Q"), 
               names_to = "Question", 
               values_to = "Value") %>% 
  # reorder "pre" and "post" for plotting
  mutate(Survey_Type = as_factor(Survey_Type)) %>% 
  # pipe into a histogram
  ggplot(aes(Value)) +
  geom_histogram(aes(fill = Survey_Type), 
                 position = "identity", alpha = 0.5, bins = 5) +
  facet_wrap(~ Question, scales = "free_x") +
  labs(x = "Score",
       y = "Frequency",
       fill = "Survey") +
  theme_bw()
```

## Run Linear Mixed Effects Models

### Question 39

Run a model for Q39 responses as a function of the survey (pre vs. post). Include student, semester, and university as random effects.

```{r}
model_Q39 <- lmer(Q39 ~ Survey_Type + (1 | Term) + (1 | School) + (1 | StudentID), 
              data = data)
broom.mixed::tidy(model_Q39)
```

Semester (`Term`) has a variance of 0, which is causing singularity. Let's remove that random effect.

```{r}
model_Q39 <- lmer(Q39 ~ Survey_Type + (1 | School) + (1 | StudentID), 
              data = data)
(output_Q39 <- broom.mixed::tidy(model_Q39))
```

This removes warning about singularity, which is great.

\| The p-value is `r round(output_Q39$p.value[2], 2)`, meaning no significant difference.

Let's check the model performance. Q39 answers do not have a normal distribution.

```{r}
resid_vals <- residuals(model_Q39)
qqnorm(resid_vals); qqline(resid_vals)
hist(resid_vals, main = "Histogram of Residuals", xlab = "Residuals")
```

Residuals aren't great, but aren't horrible, either?

Out of curiosity, what happens when we include the school as a fixed effect? We actually have reason to believe this might make more sense--one class was more computationally focused while another class was more statistically focused

```{r}
model_Q39 <- lmer(Q39 ~ Survey_Type + School + (1 | StudentID), 
              data = data)
broom.mixed::tidy(model_Q39)
```

This does not change out qualitative result at all.

### Question 15

Let's run the analysis with Question 15. For every question, the variance of semester ends up being 0, so I will pre-emptively remove it.

```{r}
model_Q15 <- lmer(Q15 ~ Survey_Type + (1 | StudentID) + (1 | School), 
                  data = data)
broom.mixed::tidy(model_Q15)
```

Check the residuals

```{r}
resid_vals <- residuals(model_Q15)
qqnorm(resid_vals); qqline(resid_vals)
hist(resid_vals, main = "Histogram of Residuals", xlab = "Residuals")
```

### Question 17_1

Let's run the analysis with Question 17_1.

```{r}
model_Q17_1 <- lmer(Q17_1 ~ Survey_Type + (1 | StudentID) + (1 | School), 
                  data = data)
broom.mixed::tidy(model_Q17_1)
```

Check the residuals

```{r}
resid_vals <- residuals(model_Q17_1)
qqnorm(resid_vals); qqline(resid_vals)
hist(resid_vals, main = "Histogram of Residuals", xlab = "Residuals")
```

### Question 17_2

```{r}
model_Q17_2 <- lmer(Q17_2 ~ Survey_Type + (1 | StudentID) + (1 | School), 
                  data = data)
broom.mixed::tidy(model_Q17_2)
```

Check the residuals

```{r}
resid_vals <- residuals(model_Q17_2)
qqnorm(resid_vals); qqline(resid_vals)
hist(resid_vals, main = "Histogram of Residuals", xlab = "Residuals")
```

### Question 17_3

Let's run the analysis with Question 17_3.

```{r}
model_Q17_3 <- lmer(Q17_3 ~ Survey_Type + (1 | StudentID) + (1 | School), 
                  data = data)
broom.mixed::tidy(model_Q17_3)
```

Check the residuals

```{r}
resid_vals <- residuals(model_Q17_3)
qqnorm(resid_vals); qqline(resid_vals)
hist(resid_vals, main = "Histogram of Residuals", xlab = "Residuals")
```

### Question 17_4

Let's run the analysis with Question 17_4.

```{r}
model_Q17_4 <- lmer(Q17_4 ~ Survey_Type + (1 | StudentID) + (1 | School), 
                  data = data)
summary(model_Q17_4)
```

Check the residuals

```{r}
resid_vals <- residuals(model_Q17_4)
qqnorm(resid_vals); qqline(resid_vals)
hist(resid_vals, main = "Histogram of Residuals", xlab = "Residuals")
```

TO DO:

-   correct p-values for multiple analyses

-   consider running analyses on absolute values–students learn if they like DS or really don't like DS, which is also fine!
