---
title: "Plots for Poster"
author: "EKB"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data in the Wild: Fall 2023

This document is for making plots based on pre- and post-surveys from the Fall 2023 iteration of Data in the Wild.

### Packages and Data

Load `tidyverse`
```{r}
library(tidyverse)
library(broom)
```

Data were compiled and sent over by Jill Williams

```{r}
data <- read_csv("../data/Fall_2023_PrePostMatched_DIW_Deidentified.csv", skip = 2) %>% 
  rename(Q39_pre = `Q39...3`, Q39_post = `Q39...4`,
         Q15_pre = `Q15...5`, Q15_post = `Q15...6`,
         Q17_1_pre = `Q17_1...7`, Q17_1_post = `Q17_1...8`,
         Q17_2_pre = `Q17_2...9`, Q17_2_post = `Q17_2...10`,
         Q17_3_pre = `Q17_3...11`, Q17_3_post = `Q17_3...12`,
         Q17_4_pre = `Q17_4...13`, Q17_4_post = `Q17_4...14`) %>% 
  select(Q39_pre:Q17_4_post) %>% 
  drop_na()
glimpse(data)
```
Pull out the questions and save for later.

```{r}
questions <- data %>% slice(1:2)
questions

data <- data %>% 
  slice(-1) %>% 
  mutate(across(everything(), as.numeric))
head(data)
```

Rearrange data so each question has pre and post values in the same column and create a new column with whether the value is pre or post

```{r}
data_pre <- data %>% 
  select(contains("pre")) %>% 
  mutate(survey = "Pre") %>% 
  rename_with(~ str_remove(., "_pre"))

data_post <- data %>% 
  select(contains("post")) %>% 
  mutate(survey = "Post") %>% 
  rename_with(~ str_remove(., "_post"))

data_long <- bind_rows(data_pre, data_post) %>% 
  mutate(survey = factor(survey, levels = c("Pre", "Post"))) %>% 
  drop_na()
head(data_long)
```
### t-tests

Series of quick t-tests to get p-values

```{r}
ttest_results <- data_long %>% 
  pivot_longer(Q39:Q17_4, names_to = "question", values_to = "response_num") %>% 
  group_by(question) %>% 
  nest() %>% 
  mutate(ttests = map(.x=data, ~tidy(pairwise.t.test(.x$response_num, .x$survey, paired = TRUE, p.adjust.method = "none")))) %>%  
  unnest(ttests) %>% 
  select(-data)
ttest_results

#write_csv(ttest_results, "../data/ttest_results.csv")
```

### Plots

Make plots for each question.

Bar plots and river plots?

Bar plots should have the average pre vs. post values with sd values

```{r}
summary_stats <- data_long %>% 
  group_by(survey) %>% 
  summarise(across(where(is.numeric), 
                   list(mean = ~ mean(., na.rm = TRUE),
                        sd = ~ sd(., na.rm = TRUE)))) 
summary_stats

summary_by_question <- summary_stats %>% 
  pivot_longer(contains("Q"), names_to = "question", values_to = "value") %>% 
  # create column for statistic type
    separate(question, c("question", "stat"), extra = "merge", fill = "right") %>% 
    separate(stat, c("temp", "stat"), fill = "left") %>% 
    unite("question", question, temp, na.rm = T) %>% 
  pivot_wider(names_from = question)
head(summary_by_question)
```

Stacked bar plots require proportions

```{r}
prop_data <- data_long %>% 
  pivot_longer(Q39:Q17_4, names_to = "question", values_to = "response_num") %>% 
  group_by(survey, question) %>% 
  mutate(total = n()) %>% 
  group_by(survey, question, response_num) %>% 
  mutate(response_count = n(),
         prop = response_count / total,
         response_word = case_when(response_num == 5 ~ "Strongly Agree",
                         response_num == 4 ~ "Somewhat Agree",
                         response_num == 3 ~ "Neither Agree nor Disagree",
                         response_num == 2 ~ "Somewhat Disagree",
                         response_num == 1 ~ "Strongly Disagree"),
         response_word = factor(response_word, levels = c("Strongly Agree", "Somewhat Agree", "Neither Agree nor Disagree", "Somewhat Disagree", "Strongly Disagree"))) %>% 
  distinct()
```


#### Question 15

Means

```{r}
p_value <- pull(filter(ttest_results, question == "Q15"), p.value)

ggplot(summary_stats, aes(x = survey, y = Q15_mean, color = survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Q15_mean - Q15_sd, ymax = Q15_mean + Q15_sd), width = 0.2) +
  geom_text(aes(label = round(Q15_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.1, vjust = 1.1, size = 6, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("../plots/Q15_means.png", height = 3.5, width = 4)
```

Stacked Bar Plot

```{r}
Q15 <- filter(prop_data, question == "Q15")

ggplot(Q15, aes(x = survey, y = prop, fill = response_word)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "left")
ggsave("../plots/Q15_stacked.png", width = 6, height = 6)
```


#### Question 17_1

Means

```{r}
p_value <- pull(filter(ttest_results, question == "Q17_1"), p.value)

ggplot(summary_stats, aes(x = survey, y = Q17_1_mean, color = survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Q17_1_mean - Q17_1_sd, ymax = Q17_1_mean + Q17_1_sd), width = 0.2) +
  geom_text(aes(label = round(Q17_1_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.1, vjust = 1.1, size = 6, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("../plots/Q17_1_means.png", height = 3.5, width = 4)
```

Stacked Bars

```{r}
Q17_1 <- filter(prop_data, question == "Q17_1")

ggplot(Q17_1, aes(x = survey, y = prop, fill = response_word)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "none")

ggsave("../plots/Q17_1_stacked.png", width = 4.5, height = 6)
```

#### Question 17_2

Means

```{r}
p_value <- pull(filter(ttest_results, question == "Q17_2"), p.value)

ggplot(summary_stats, aes(x = survey, y = Q17_2_mean, color = survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Q17_2_mean - Q17_2_sd, ymax = Q17_2_mean + Q17_2_sd), width = 0.2) +
  geom_text(aes(label = round(Q17_2_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
 annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.1, vjust = 1.1, size = 6, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("../plots/Q17_2_means.png", height = 3.5, width = 4)
```

Stacked Bars

```{r}
Q17_2 <- filter(prop_data, question == "Q17_2")

ggplot(Q17_2, aes(x = survey, y = prop, fill = response_word)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "none")

ggsave("../plots/Q17_2_stacked.png", width = 4.5, height = 6)
```

#### Question 17_3

Means

```{r}
p_value <- pull(filter(ttest_results, question == "Q17_3"), p.value)

ggplot(summary_stats, aes(x = survey, y = Q17_3_mean, color = survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Q17_3_mean - Q17_3_sd, ymax = Q17_3_mean + Q17_3_sd), width = 0.2) +
  geom_text(aes(label = round(Q17_3_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.1, vjust = 1.1, size = 6, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("../plots/Q17_3_means.png", height = 3.5, width = 4)
```

Stacked Bars

```{r}
Q17_3 <- filter(prop_data, question == "Q17_3")

ggplot(Q17_3, aes(x = survey, y = prop, fill = response_word)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "none")

ggsave("../plots/Q17_3_stacked.png", width = 4.5, height = 6)
```

#### Question 17_4

Means

```{r}
p_value <- pull(filter(ttest_results, question == "Q17_4"), p.value)

ggplot(summary_stats, aes(x = survey, y = Q17_4_mean, color = survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Q17_4_mean - Q17_4_sd, ymax = Q17_4_mean + Q17_4_sd), width = 0.2) +
  geom_text(aes(label = round(Q17_4_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.1, vjust = 1.1, size = 6, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("../plots/Q17_4_means.png", height = 3.5, width = 4)
```

Stacked Bars

```{r}
Q17_4 <- filter(prop_data, question == "Q17_4")

ggplot(Q17_4, aes(x = survey, y = prop, fill = response_word)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "none")

ggsave("../plots/Q17_4_stacked.png", width = 4.5, height = 6)
```



