---
title: "DITW for SABER West"
author: "Ellen Bledsoe"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# SABER West Analyses

## Data Wrangling

Load the `tidyverse`
```{r}
library(tidyverse)
library(broom)
```

Read in the datasets: one for pre-surveys and one for post-surveys.
```{r}
# remove first 2 rows
pre <- read_csv("data/presurveys.csv") %>% 
  slice(-c(1:2)) %>% 
  rename(Institution = Q1, Term = Q42) 

post <- read_csv("data/postsurveys.csv") %>% 
  slice(-c(1:2)) %>% 
  select(-1) %>% 
  rename(Institution = Q1, Term = Q70)
```
Join datasets
```{r}
# join datasets and create a column for pre- vs. post-survey
all <- bind_rows(pre, post, .id = "Survey") %>% 
  drop_na() %>% 
  mutate(Survey = if_else(Survey == 1, "Pre", "Post"),
         Survey = factor(Survey, levels = c("Pre", "Post")))
```

Convert Data to Long Form
```{r}
# long-form data and add numbers for Likert scale
all <- all %>% 
  pivot_longer(cols = Q15:Q17_4, names_to = "Question", values_to = "Response_Words") %>% 
  mutate(Response_Numbers = if_else(Question == "Q15",
                                    case_when(Response_Words == "Extremely interested" ~ 5,
                                              Response_Words == "Very interested" ~ 4,
                                              Response_Words == "Moderately interested" ~ 3,
                                              Response_Words == "Slightly interested" ~ 2,
                                              Response_Words == "Not interested at all" ~ 1),
                                    case_when(Response_Words == "Strongly agree" ~ 5,
                                              Response_Words == "Somewhat agree" ~ 4,
                                              Response_Words == "Neither agree nor disagree" ~ 3,
                                              Response_Words == "Somewhat disagree" ~ 2,
                                              Response_Words == "Strongly disagree" ~ 1)),
         Response_Words = case_when(Response_Words == "Extremely interested" ~ "Extremely Interested",
                                    Response_Words == "Very interested" ~ "Very Interested",
                                    Response_Words == "Moderately interested" ~ "Moderately Interested",
                                    Response_Words == "Slightly interested" ~ "Slightly Interested",
                                    Response_Words == "Not interested at all" ~ "Not Interested At All",
                                    Response_Words == "Strongly agree" ~ "Strongly Agree",
                                    Response_Words == "Somewhat agree" ~ "Somewhat Agree",
                                    Response_Words == "Neither agree nor disagree" ~ "Neither Agree nor Disagree",
                                    Response_Words == "Somewhat disagree" ~ "Somewhat Disagree",
                                    Response_Words == "Strongly disagree" ~ "Strongly Disagree"))

```

Calculate proportions for plotting

```{r}
# create grid with every possible combo of Survey, Question and Response
full_grid <- expand_grid(Survey = unique(all$Survey),
                         Question = unique(all$Question),
                         Response_Words = unique(all$Response_Words)) %>% 
  filter((Question == "Q15" & str_detect(Response_Words, "Interest")) | 
            (Question != "Q15" & !str_detect(Response_Words, "Interest")))

# calculate proportions for each answer
prop_data <- all %>% 
  # create column for number of total responses
  group_by(Survey, Question) %>% 
  mutate(Total = n()) %>% 
  # create columns for number of responses per category and proportion
  group_by(Survey, Question, Response_Words) %>% 
  mutate(Response_Count = n(),
         Prop = Response_Count / Total) %>% 
  # join with full_grid dataframe to include rows for missing response combos (did I need to?)
  full_join(., full_grid) %>% 
  # arrange them to more easily see responses
  distinct(across(.cols = c(Response_Words, Response_Numbers, Prop, Response_Count, Total))) %>% 
  arrange(Survey, Question, Response_Numbers)
```

Perform t-tests

```{r}
t_tests <- all %>% 
  group_by(Question) %>% 
  mutate(Survey = as.character(Survey)) %>% 
  nest() %>% 
  mutate(ttests = map(.x = data, ~tidy(t.test(.x$Response_Numbers ~ .x$Survey)))) %>%  
  unnest(ttests) %>% 
  select(-data)
```
Make plots for each question.

Bar plots and river plots?

Bar plots should have the average pre vs. post values with sd values

```{r}
summary_stats <- all %>% 
  group_by(Question, Survey) %>% 
  summarise(across(where(is.numeric), 
                   list(mean = ~ mean(., na.rm = TRUE),
                        sd = ~ sd(., na.rm = TRUE)))) 
summary_stats
```

## Plotting

### Proportion Data

Plotting Question 15
```{r}
Q15 <- filter(prop_data, Question == "Q15") %>% 
  mutate(Response_Words = factor(Response_Words, levels = c("Extremely Interested", "Very Interested", "Moderately Interested", "Slightly Interested", "Not Interested At All"))) 

ggplot(Q15, aes(x = Survey, y = Prop, fill = Response_Words)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "left")

ggsave("plots/Q15_stacked.png", width = 6, height = 6)
```

Add factor levels to Q17s
```{r}
prop_data <- filter(prop_data, Question != "Q15") %>% 
  mutate(Response_Words = factor(Response_Words, levels = c("Strongly Agree", "Somewhat Agree", "Neither Agree nor Disagree", "Somewhat Disagree", "Strongly Disagree")))
```


Q17_1

```{r}
Q17_1 <- filter(prop_data, Question == "Q17_1")

ggplot(Q17_1, aes(x = Survey, y = Prop, fill = Response_Words)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "left")

ggsave("plots/Q17_1_stacked.png", width = 6.5, height = 6)
```

Q17_2

```{r}
Q17_2 <- filter(prop_data, Question == "Q17_2")

ggplot(Q17_2, aes(x = Survey, y = Prop, fill = Response_Words)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "none")

ggsave("plots/Q17_2_stacked.png", width = 4.5, height = 6)
```
Q17_3

```{r}
Q17_3 <- filter(prop_data, Question == "Q17_3")

ggplot(Q17_3, aes(x = Survey, y = Prop, fill = Response_Words)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "none")

ggsave("plots/Q17_3_stacked.png", width = 4.5, height = 6)
```

Q17_4

```{r}
Q17_4 <- filter(prop_data, Question == "Q17_4")

ggplot(Q17_4, aes(x = Survey, y = Prop, fill = Response_Words)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(direction = 1) +
  labs(x = "Survey",
       fill = "Response") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14),
        legend.position = "none")

ggsave("plots/Q17_4_stacked.png", width = 4.5, height = 6)
```

### Means

Q15

```{r}
p_value <- pull(filter(t_tests, Question == "Q15"), p.value)
Q15 <- filter(summary_stats, Question == "Q15")

ggplot(Q15, aes(x = Survey, y = Response_Numbers_mean, color = Survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Response_Numbers_mean - Response_Numbers_sd, 
                    ymax = Response_Numbers_mean + Response_Numbers_sd), width = 0.2) +
  geom_text(aes(label = round(Response_Numbers_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = "p < 0.01", hjust = 1.25, vjust = 16, size = 5, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("plots/Q15_means.png", height = 3, width = 4)
```

Q17_1

```{r}
p_value <- pull(filter(t_tests, Question == "Q17_1"), p.value)
Q17_1 <- filter(summary_stats, Question == "Q17_1")

ggplot(Q17_1, aes(x = Survey, y = Response_Numbers_mean, color = Survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Response_Numbers_mean - Response_Numbers_sd, 
                    ymax = Response_Numbers_mean + Response_Numbers_sd), width = 0.2) +
  geom_text(aes(label = round(Response_Numbers_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.25, vjust = 16, size = 5, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("plots/Q17_1_means.png", height = 3, width = 4)
```

Q17_2

```{r}
p_value <- pull(filter(t_tests, Question == "Q17_2"), p.value)
Q17_2 <- filter(summary_stats, Question == "Q17_2")

ggplot(Q17_2, aes(x = Survey, y = Response_Numbers_mean, color = Survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Response_Numbers_mean - Response_Numbers_sd, 
                    ymax = Response_Numbers_mean + Response_Numbers_sd), width = 0.2) +
  geom_text(aes(label = round(Response_Numbers_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
 annotate("text", x = Inf, y = Inf, label = "p < 0.01", hjust = 1.25, vjust = 16, size = 5, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("plots/Q17_2_means.png", height = 3, width = 4)
```

Q17_3

```{r}
p_value <- pull(filter(t_tests, Question == "Q17_3"), p.value)
Q17_3 <- filter(summary_stats, Question == "Q17_3")

ggplot(Q17_3, aes(x = Survey, y = Response_Numbers_mean, color = Survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Response_Numbers_mean - Response_Numbers_sd, 
                    ymax = Response_Numbers_mean + Response_Numbers_sd), width = 0.2) +
  geom_text(aes(label = round(Response_Numbers_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.25, vjust = 16, size = 5, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("plots/Q17_3_means.png", height = 3, width = 4)
```
Q17_4

```{r}
p_value <- pull(filter(t_tests, Question == "Q17_4"), p.value)
Q17_4 <- filter(summary_stats, Question == "Q17_4")

ggplot(Q17_4, aes(x = Survey, y = Response_Numbers_mean, color = Survey)) +
  geom_point(size = 5) +
  geom_errorbar(aes(ymin = Response_Numbers_mean - Response_Numbers_sd, 
                    ymax = Response_Numbers_mean + Response_Numbers_sd), width = 0.2) +
  geom_text(aes(label = round(Response_Numbers_mean, 1)), hjust = -0.75, color = "black", fontface = "bold", size = 8) +
  annotate("text", x = Inf, y = Inf, label = paste("p = ", round(p_value, 2)), hjust = 1.25, vjust = 16, size = 5, fontface = "italic") +
  scale_y_continuous(limits = c(0, 5.5),
                     breaks = seq(0, 5, by = 1)) +
  scale_color_viridis_d(direction = 1, begin = 0, end = 0) +
  labs(x = "Survey",
       fill = "Survey") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 14))

ggsave("plots/Q17_4_means.png", height = 3, width = 4)
```

